---
title: "Can you get a shiny server on DO without leaving RStudio?"
author: "Rob Cavanaugh"
date: "2/12/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

When I first created a shiny server, I followed the instructions laid out 
here [link] and here [link] which was a good learning exercise, required a lot of
trial and error and was pretty time consuming. I suspect that the process is
also a bit too much for many R users. My wife recently started working for DO and 
mentioned that there is a new add-on that will make installing a shiny app
a pretty quick process. I also saw some recent updates to the 
analogsea package [here] and started getting curious - is it now possible to 
create a shiny server on a DO droplet without leaving Rstudio (or at least without
using terminal). Unfortunately, I couldn't find any guides or posts on how
to do this, so I documented my process...

Note that I figured this out mostly from trial and error...there isn't a whole lot
of documentation for doing this. Aspects of this could be wrong. Also DO is not expensive,
and you start with lots of credits, but its not free like the free plan of shinyapps.io. 

As a spoiler, its not possible to do everything by the command line. But these
steps should save a substantial amount of time relative to existing guides that 
have been helpful to me: 

- https://www.marinedatascience.co/blog/2019/04/28/run-shiny-server-on-your-own-digitalocean-droplet-part-1/#step-4-add-more-security-by-changing-the-default-port-optional

- https://deanattali.com/2015/05/09/setup-rstudio-shiny-server-digital-ocean/



```{r, eval = T, warning = F, message = F}
# packages
library(here)
library(tidyverse)
library(analogsea)
```

1. Create a [Digital Ocean](https://m.do.co/c/ad9660bd69ac) account. (Note, this is my referral link...)

2. SSH link computer to DO: https://docs.digitalocean.com/products/droplets/how-to/add-ssh-keys/
Unfortunately, I already had to leave Rstudio. but you can connect your ssh key to DO once you create it by using `analogsea::key_create("my_ssh_key", readLines("~/.ssh/id_rsa.pub"))` where `"my_ssh_key"` is whatever you'd like to name your ssh key. 

3. Set an API-key for analogsea. You can create the key here: https://cloud.digitalocean.com/account/api/

```{r}
Sys.setenv(DO_PAT = "KEY_GOES_HERE")
```

4. I already had a droplet so I wanted to check that I could see it. If you don't 
have a droplet yet...this will probably return NULL or an error. 

```{r, eval = T}
droplets()
```

5. Create a droplet (virtual computer). This required the most trial and error
on my part. I tried using `docklet_create()` and then `docklet_shinyserver()`
which successfully created a containerized virual machine with shiny-server
pre-installed. However, installing packages and otherwise working within the
container is a little over my paygrade and I ran into too many errors using
the analogsea package. Similar issues with `docklet_rstudio()` so I abandoned
that tack. 

Then I tried just creating a regular droplet, however, this required then going
through the same steps I did last time I manually set up a shiny server. Reading
the documentation for `droplet_create` revealed this tidbit: 

> image: (character/numeric) The image ID of a public or private image, or the unique slug identifier for a public image. This image will be the base image for your droplet. See images() for a complete list. Use rstudio-20-04 for a DigitalOcean Marketplace image with R and Tidyverse readily available. Default: ubuntu-18-04-x64

This is what I really wanted, the `rstudio-20-04` image refers to [this](https://marketplace.digitalocean.com/apps/rstudio) specific add-on which
includes *almost* everything I wanted to do pre-configured. I gave the new droplet
a name `"shinyDO"` but this could be anything. 

```{r}
droplet_create(name = "shinyDO", image = "rstudio-20-04")
```

After some messages in the console, we can check that the droplet was created
successfully:

```{r}
droplets()
```

Note: When you create a droplet using this function, you can save its information
 by running, for example, `d <- droplet_create(...)` but I liked rather being
 explicit about creating this object. 

```{r}
d <- droplets()$shinyDO
```

6. Check that rstudio server and shiny server are working:

- Go to the droplet `$shinyDO` IP which in this case is 137.184.113.199:8787 to see 
the rstudio instance 
- 137.184.113.199:3838 to see the shiny server example page.

7. There are some extra steps that are recommended on the [add-on page](https://marketplace.digitalocean.com/apps/rstudio)

- Change the password to rstudio server. I wasn't able to log in with 
the supposed defaults rstudio/rstudio, so I skipped this step and came back to
it. 
- Create a separate admin account

```{r}
# create the new user and password. 
d %>% ubuntu_create_user(user = "rob", password = "password1234")

# give this user admin rights
d %>% droplet_ssh("usermod -aG sudo rob")

# delete the rstudio user
# this command failed, saying the user
# rstudio does not exist. 
# now I understand why the initial logging in didn't work...
# But is there some default user that I need to change
# the password for?

# d %>% droplet_ssh("deluser rstudio")
```

Note here the use of the function `droplet_ssh()` which executes command line code
in the app as the current user (usually root). This could be done also by running
`ssh root@IPaddress` in terminal or using the DO command prompt as referenced
in the above page. I'm using `droplet_ssh()` here because the goal was to 
stay in rstudio as much as possible. But some things may not work if
they require additional input. For these, its better to use the other options.

7. Install extra r packages

- I need these for the example app I'm going to upload.

```{r}
d %>% install_r_package(c("glue", "shinyMobile", "rtide", "jsonlite", "lubridate", "httr", "ggpmisc"))
```

8. Upload a shiny app to the correct location. 

According to the shiny server documentation, shiny apps are served off of the folder
`/srv/shiny-server`. Droplet upload appears to upload all files within a folder
so I added the name of the folder I wanted to upload files under to `shiny-server/`. 
This seemed to work fine. Note that the name you give this folder will be the extension
of where you find the app. So now, IPaddress:3838/portland-tide is where the app
can be accessed. 

```{r}
d %>% droplet_upload(here::here(), "/srv/shiny-server/portland-tide")
```

Other useful commands, using the `droplet_ssh()` command.

Just in case, how to remove an app

```{r}
d %>% droplet_ssh("sudo rm -rf /srv/shiny-server/portland-tide")
```

Or check the app log files

```{r}
d %>% droplet_ssh("tail -f -n100 /var/log/shiny-server.log")
```

Note that there are other commands that might be included in {analogsea} so that
you don't need to use `droplet_ssh()`.

# Domain stuff

1. Buy a domain
2. Configure it to digital ocean
3. test that you can go to your domain name (domain.com) and see the
nginx default page.


Securing your website:
The following steps you will need to do in terminal via ssh or the DO terminal.

https://www.digitalocean.com/community/tutorials/how-to-secure-nginx-with-let-s-encrypt-on-ubuntu-20-04

Then, in the terminal, add your domain to the server name line. you can edit it with
`sudo nano /etc/nginx/sites-available/default`.

Change `server_name _;` to `server_name domain.com;`

Closer the file. then run `sudo service nginx restart`



